Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-18 forecast/paper/paper.tex
fix author name typo
Update sig statement

% Please add a significance statement to explain the relevance of your work
\significancestatement{
  TODO % For reference, below is the indicators paper one.

  % To study the COVID-19 pandemic, its effects on society,
  % and measures for reducing its spread, researchers need detailed data on the
  % course of the pandemic. Standard public health data streams suffer
  % inconsistent reporting and frequent, unexpected revisions. They also miss
  % other aspects of a population's behavior, worthy of consideration. We
  % present an open database of COVID signals in the United States, measured at
  % the county level and updated daily. These include traditionally reported
  % COVID cases and deaths, and many others: signals based on measures of
  % mobility, social distancing, internet search trends, self-reported symptoms,
  % and patterns COVID-related activity in de-identified medical insurance 
  % claims. The database provides all signals in a common, easy-to-use format,
  % empowering both  public health research and operational decision-making. 
}

@@

Figure~\ref{fig:forecast} (right panel) displays the results for forecasting, stratified by
ahead value and averaged over all HRRs and forecast dates (June 9--December 31,
2020).  Shown is the average WIS for each forecast model divided by the averaged
WIS from a baseline model, which is basically a flat-line forecaster (its
median forecast for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted
quantiles defined around this flat line based on historical variation). This is
the same baseline model as that used in the COVID-19 Forecast Hub.  Here, we  
divide by the average WIS of the baseline model in order to put the y-axis on an 
interpretable scale.  We can see that all curves are below 1, which means
(smaller WIS is better) that all of the models, including AR, outperform the
baseline on average over the forecasting period.  On the other hand, the models 
deliver at best an improvement of about 20\% in mean WIS over the baseline
model, with this gap narrowing to only 10\% at the largest ahead values,
illustrating the difficulty of the forecasting problem.

@@

% \begin{figure}[t]
%   \includegraphics[width=\columnwidth]{fig/fcast-1.pdf}
%   \caption{Average WIS for each forecast model, over all forecast dates and all
%     HRRs, divided by the average WIS achieved by a baseline model (a
%     probabilistic version of the flat-line forecaster).  Here and in all figures
%   we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.}  
%   \label{fig:forecast}
% \end{figure}

Figure~\ref{fig:forecast} (left panel) displays the results for hotspot prediction, again
stratified by ahead value and averaged over all HRRs and prediction dates (June 
16--December 31, 2020).  We can see many similarities to the forecasting
results (now larger AUC is better); for example, CHNG-COVID and DV-CLI offer the
biggest improvement over AR, and all models including AR degrade in performance 
towards the baseline (this is now just a classifier based on random guessing,
which achieves an AUC of 0.5) as the ahead values grow.  A clear difference,
however, is that the CHNG-CLI model performs quite well in hotspot prediction,
close to the best-performing indicator models for many of the ahead values. This
may be because volatility in the CHNG-CLI indicator plays less of a role in the
associated logistic model's predicted probabilities (in general, a sigmoid
function can absorb a lot of the variability in its input).

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
  \caption{Left: Average WIS for each forecast model, over all forecast dates and all
    HRRs, divided by the average WIS achieved by a baseline model (a
    probabilistic version of the flat-line forecaster).  Here and in all figures
    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.
    Right: Area under the curve for each hotspot prediction model, calculated over
    all prediction dates and all HRRs.
  }  
  \label{fig:forecast}
\end{figure*}

>>>>

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element of the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  Internet search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

  % TODO % For reference, below is the indicators paper one.
  % To study the COVID-19 pandemic, its effects on society,
  % and measures for reducing its spread, researchers need detailed data on the
  % course of the pandemic. Standard public health data streams suffer
  % inconsistent reporting and frequent, unexpected revisions. They also miss
  % other aspects of a population's behavior, worthy of consideration. We
  % present an open database of COVID signals in the United States, measured at
  % the county level and updated daily. These include traditionally reported
  % COVID cases and deaths, and many others: signals based on measures of
  % mobility, social distancing, internet search trends, self-reported symptoms,
  % and patterns COVID-related activity in de-identified medical insurance 
  % claims. The database provides all signals in a common, easy-to-use format,
  % empowering both  public health research and operational decision-making. 
%}

@@

Figure~\ref{fig:forecast} (left panel) displays evaluation results for
forecasting, stratified by ahead value and averaged over all HRRs and forecast
dates.  Shown is the average WIS for each forecast model divided by that from a
baseline model, which is basically a flat-line forecaster (its median forecast
for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
around this based on historical variation). This is the same baseline model as
in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
mean WIS so that it gets put on an interpretable, unitless scale.  In the
figure, we can see that all curves are below 1, which means (smaller WIS is
better) that all of the models, including AR, outperform the baseline on average
over the forecasting period.  On the other hand, the models deliver at best an
improvement of about 20\% in average WIS over the baseline model, with this gap
narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
of the forecasting problem.

@@

Figure~\ref{fig:forecast} (right panel) displays the results for hotspot
prediction, again stratified by ahead value and averaged over all HRRs and
prediction dates.  We can see many similarities to the forecasting results (now
larger AUC is better); for example, CHNG-COVID and DV-CLI offer the biggest
improvement over AR, and all models including AR degrade in performance towards
the baseline (this is now just a classifier based on random guessing, which
achieves an AUC of 0.5) as the ahead values grow, illustrating the difficulty of
the hotspot prediction problem A clear difference, however, is that the CHNG-CLI
model performs quite well in hotspot prediction, close to the best-performing
indicator models for many of the ahead values. This may be because volatility in
the CHNG-CLI indicator plays less of a role in the associated logistic model's
predicted probabilities (in general, a sigmoid function can absorb a lot of the
variability in its input).

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
  \caption{Left: average WIS for each forecast model, over all forecast dates
    and all HRRs, divided by the average WIS achieved by a baseline model (a
    probabilistic version of the flat-line forecaster).  Right: area under the
    curve for each hotspot prediction model, calculated over all prediction
    dates and all HRRs.  Here and in all figures we abbreviate
    CTIS-CLI-in-community by CTIS-CLIIC.  
  }  
  \label{fig:forecast}
\end{figure*}
